---
title: "Single-taxon indicator (trim)"
author: "William K. Morris"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Single-taxon indicator (trim)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path = "figure/")
knitr::opts_knit$set(base.dir = "pkg/vignettes/")
```

This documents outlines how a single-taxon indicator (_Anthus pratensis_
breeding on farmland) can be calculated in R using the TRIM method.

## Load packages

The following packages are required. All packages are available on
[CRAN](https://cran.r-project.org/) apart from `{fbi}` which can be installed
from GitHub.

```{r pkgs, message = FALSE}
library(fbi)
library(finbif)
library(ggplot2)
library(lubridate)
library(rtrim)
```

## Survey data

These five fields are required for the survey data.

```{r survey-select}
select <- c("document_id", "location_id", "year", "month", "day")
```

These filters restrict the survey data to the "Point count" and "Line transect"
bird monitoring datasets from 1979 onwards at sites labelled "farmland" and
where the selected data fields have no missing data.

```{r survey-filter}
filter <- list(
  location_tag = "farmland",
  collection = c(
    "Point counts of breeding terrestrial birds",
    "Line transect censuses of breeding birds"
  ),
  date_range_ymd = c("1979-01-01", ""),
  has_value = select
)
```

The survey data can now downloaded from FinBIF.

```{r surveys}
surveys <- finbif_occurrence(
  filter = filter,
  select = select,
  aggregate = "events",
  aggregate_counts = FALSE,
  n = "all",
  quiet = TRUE
)
```

Two processing functions are applied to the survey data to first limit each site
to the first survey of the year and then limit the surveys to sites where at
least two years have been surveyed.

```{r process-survyes}
surveys <- pick_first_survey_in_year(surveys)

surveys <- require_two_years(surveys)
```

## Count data

Count data only requires two fields to be selected: the survey identifier
(`document_id`) and the measure of abundance (in this case the number of
breeding pairs: `pair_abundance`).

```{r count-select}
select <- c("document_id", abundance = "pair_abundance")
```

The count data requires the same filters as the survey data (though the filter
`has_value` needs to be redefined).

```{r count-filter}
filter[["has_value"]] <- select
```

The count data for _Anthus pratensis_ (Meadow pipit) can now be downloaded from
FinBIF.

```{r counts}
counts <- finbif_occurrence(
  taxa = "Anthus pratensis",
  filter = filter,
  select = select,
  n = "all",
  quiet = TRUE
)
```

Three processing functions are applied to the count data to: infill the
count data with zero occurrences using the survey data; sum over the counts
for each site-year combination; and remove all sites where the number of
breeding pairs was zero on every occasion.

```{r process-counts}
counts <- zero_fill(counts, surveys)

counts <- sum_by_event(counts)

counts <- remove_all_zero_locations(counts)
```

# Fit TRIM Model

A TRIM model is used to estimate the change in abundance over time.

```{r model}
model <- trim(abundance ~ location_id + year, counts)
```

# Create Index

An index of change in relative abundance is created by setting the base year to
the year 2000.

```{r index}
index <- index(model, base = 2000)
```

```{r plot-single-taxon-trim}
#| fig.cap: Relative abundance of Meadow pipit breeding pairs on Farmland in Finland (calculated using rTRIM)

ggplot(index) +
aes(
  x = parse_date_time(time, "Y"), y = imputed,
  ymin = imputed - se_imp, ymax = imputed + se_imp
) +
geom_ribbon(alpha = .2) +
geom_line() +
ylab(NULL) +
xlab(NULL) +
theme_minimal()
```
